Hydrodynamics and rheology of adhesive vesicle doublets

### 2019-02-05 vesicle_code/docs/adhesion/adhesion.tex
added a caption
first pass through intro.  Expanded journal names in .bib and fixed other bibtex isues

\title{Hydrodynamics and rheology of weakly adhesive vesicle doublets}
% Force line breaks with \\
%\thanks{A footnote to the article title}%

@@

\section{Introduction}
Vesicles (closed fluid-filled phospholipid bilayer membranes) have been
widely utilized as biological cell mimics in biophysics and material
engineering \cite{}.  The ever expanding applications have encouraged
detailed experimental, theoretical and numerical investigations of
vesicle hydrodynamics in applied flow, electric and magnetic fields.
Theoretical investigations often take the form of small-deformation
analysis or spheroidal dynamics of a nearly-spherical vesicle (or
nearly-circular in two dimensions) subjected to an applied flow or under
an external forcing~\cite{Barthes-BieselRallison1981_JFM,
Misbah2006_PRL, Vlahovska2007_PRE, Finken2008_EPL,
ZhangZahnTanLin2013_PoF, Nganguia2013_PRE}.  On the other hand, various
numerical methods have been developed for simulating the transient
hydrodynamics of vesicle suspensions~\cite{BagchiJohoson2005_JBE,
Biben2005_EJP, Veerapaneni2009_JCP, SeolHuKimLai2016_JCP}.  
%Experiments are performed to study vesicle hydrodynamics under various flowing conditions or forcing configurations \cite{MaderVitkova2006_EurPhysJE, Abkarian2008_SoftMatt,DahlNarsimhanGouveia2016_SoftMatt}.

Hydrodynamics of a single vesicle in Stokes flow has been extensively
investigated. In a planar shear flow, the vesicle hydrodynamics is
characterized by excess area (reduced volume), viscosity contrast
between interior and exterior fluids, and shear rate of the imposed
far-field fluid flow. In addition a vesicle with a rigid particle inside
is also investigated as a biological mimic of a eukaryotic cell with a
nucleus that occupies nearly half of the intracellular
volume~\cite{Veerapaneni2011_PRL}. Small-deformation analysis shows that
a vesicle tank-treads in a planar shear flow for low viscosity contrast
and shear rate. At high viscosity contrast the tank-tread (TT) dynamics
transitions to tumbling (TB)~\cite{Misbah2006_PRL, Vlahovska2007_PRE},
and this leads to a transition in effective shear viscosity in the
vesicle suspension~\cite{Misbah2006_PRL,Vitkova2008_BJ} that is also
validated by direct numerical
simulations~\cite{GhigliottiBibenMisbah2010_JFM} and
experiments~\cite{KantslerSegreSteinberg2008_EPL,
ZabuskySegreDeschamps2011_PoF}.  Between TT and TB vesicle dynamics, a breathing (tremble) mode is also
observed~\cite{Misbah2006_PRL,KantslerSegreSteinberg2008_PRL,
ZhaoShaqfeh2011_JFM, SpannZhaoShaqfeh2014_PoF} to affect the effective
shear viscosity in a different way.  In an extensional flow (planar or
uniaxial), vesicle shape dynamics depends sensitively on the vesicle
reduced volume  and the elastic capillary
number~\cite{KantslerSegreSteinberg2008_PRL, ZhaoShaqfeh2013_JFM,
Narsimhan2014_JFM, DahlNarsimhanGouveia2016_SoftMatt}: Asymmetric shape
and oscillatory undulation of the vesicle membrane are two examples of
the complex vesicle hydrodynamics in an extensional flow.

%
% for example, 
%it was shown recently that vesicles may align to form chains along the direction of the externally applied DC electric field with ensuing consequences to suspension rheology \cite{wu2018pairwise}.
%
Collective hydrodynamics of vesicles is dictated by the individual
vesicle-vesicle interactions.  
In a quiescent flow,
vesicle-vesicle adhesion leads to formation of vesicle doublets \cite{Ziherl2007_PRL,ZiherlSvetina2007_PNAS} or clusters \cite{SvetinaZiherl2008_Bioelectrochemistry,FlormannAouane2017_SciReports}.
As a model for red blood cell (RBC) aggregates, an idealized model for adhesive vesicle-vesicle interactions gives rise to 
vesicle configurations similar to those observed in experiments of fibrinogen-induced RBC aggregates \cite{SvetinaZiherl2008_Bioelectrochemistry,FlormannAouane2017_SciReports}. 
Using the L-J potential for point-point interaction between two vesicle membranes without the molecular details of
adhesive interactions between RBCs, Flormann {\it et al.} found the contact zone to buckle under strong adhesion \cite{FlormannAouane2017_SciReports}. 
A direct consequence of the buckling is the
sigmoidal shape of the adhesion region, which is also observed in RBC doublets and explained in a slightly different model \cite{ZiherlSvetina2007_PNAS}. 
It remains unknown how adhesive interactions between vesicles might lead to different vesicle hydrodynamics that results in
different rheological properties of a vesicle suspension. Such studies will provide useful insight to the rheology of blood flow and how to use nano solutes to control the hypertension
by adjusting the adhesive interactions between RBCs.
%

In numerical simulations of  a vesicle suspension, the adhesive interaction between two vesicles is often 
ignored \cite{Veerapaneni2009_JCP,RahimianVeerapaneniBiros2010_JCP} 
for numerical convenience, mostly 
because a smaller time step is often needed to resolve the dynamics of the lubrication thin film between two membranes with adhesion than without.
However, adhesive interactions  between RBCs lead to different RBC clusters that are expected to alter the rheological properties of the RBC suspension \cite{NeuMeiselman2002_BJ,SvetinaZiherl2008_Bioelectrochemistry}.
%
Hydrodynamic simulations of adhesive vesicles in a quiescent flow have revealed physical insights to the equilibrium shapes of RBC aggregates in both experiments 
\cite{FlormannAouane2017_SciReports} and theory \cite{ZiherlSvetina2007_PNAS}. The main goal in this work is to investigate the hydrodynamics of adhesive vesicles in flowing
conditions common in microfluidic vesicle flow, such as a planar shear flow or an extensional flow.

%
The electrostatic nature of lipid molecules (a hydrophobic tail and a hydrophilic head
with an electric dipole) complicates the interactions between a lipid
bilayer membrane and another bilayer
membrane~\cite{EvansMetcalfe1984_BJ, Book_PhysicalBasisCellAdhesion,Book_IntermolecularSurfaceForces,PerutkovaFrank-Bertoncelij2013_CSB} 
or a solid (such as a glass substrate
or a nano particle).  Adhesion between a vesicle and a substrate has been extensively studied with 
more focuses on the static equilibrium
shapes~\cite{Seifert1990_PRA, ShiFengGao2006_ActaMechSin,
LinFreund2007_IntJSolidsStructures, GruhnFrankeDimova2007_Langmuir,das2008adhesion, zhang2009phase,SteinkuhlerAgudo-Canalejo2016_BJ}
than the transient adhesion process \cite{cantat1999lift, suk-sei2001, BlountMiksisDavis2013_PRSa}.
%
Adhesive interactions between lipid membranes are essential in many biomedical, biological and biophysical
processes.  For example, vesicle adhesion is crucial to initiate membrane
fusion/fission in endocytosis and exocytosis, the transport of small
vesicles through membrane surfaces.
In the absence of an external electric field and ions in the solvents, 
it is reported that 
%such complicated electrostatic component in 
the adhesive interactions between two membranes can be well approximated by the 
L-J-type potential \cite{FlormannAouane2017_SciReports}, which consists of a long-range attraction component and a
short-range repulsion component~\cite{Book_IntermolecularSurfaceForces},
and the combination of the two gives rise to adhesion at the separation
distance that minimizes the interaction
potential~\cite{Book_IntermolecularSurfaceForces}. 
%%
%%move this somewhere else Jan30YNY
%Such adhesive
%potentials are essential in many biomedical, biological and biophysical
%processes.  For example, vesicle adhesion is essential for membrane
%fusion/fission in endocytosis and exocytosis, the transport of small
%vesicles through membrane surfaces.
%%move this somewhere else Jan30YNY

%Adhesion between bilayer membranes was first experimentally observed by
Evans and Metcalfe~\cite{EvansMetcalfe1984_BJ} used the micropipette to measure the adhesive force between two vesicles bound by strong adhesion.
They were able to  measure the
reduction in the free energy per unit area of membrane-membrane contact
formation due to the van der Waals' attraction. For the adhesive interaction between two lipid bilayer membranes in buffer solutions, a typical range for the
energy density is between $1$ to $10$ $\mu J/m^2$, 
%
%they reported
%an energy density of $\sim 10$  $\mu J/m^2$ and a separation distance of $\sim 3$  $nm$.
%
% membrane-membrane
%separation distance of $\sim 3$ $nm$, they estimated that the Hamaker
%constant is approximately $5.8\times 10^{-14}$ $ergs$, which corresponds
%to 
%
%an adhesion energy density around $1$ $\mu J/m^2$,
similar to the adhesion energy density of $3$  $\mu J/m^2$ between two RBCs \cite{FlormannAouane2017_SciReports}.
For adhesion interaction between a vesicle and a substrate  \cite{GruhnFrankeDimova2007_Langmuir}, 
the boundary between weak and strong adhesion is around $1$ $\mu J/m^2$:
the vesicle-substrate interaction is ``strong" when the adhesion energy density is larger than $1$ $\mu J/m^2$ and ``weak" when the energy
density is less than $1$ $\mu J/m^2$.
%
%weak adhesion regime 
%
%Gruhn {\it et al.} made a distinction between strong and weak adhesion regimes for a vesicle adhering on a substrate.
%Strong adhesion regime when the adhesion energy density is greater than $1$ $\mu J/m^2$,
%while weak adhesion regime when adhesion energy density is less than $1$ $\mu J/m^2$.
%
In this work we use the ratio of the adhesion energy to the bending energy for classification of adhesion strength:
The ratio of total adhesion
energy to the bending energy (bending rigidity of lipid
bilayer membranes $\sim 10^{-19}$ $J$) gives a measure of vesicle
deformation in the presence of
adhesion~\cite{RamachandranAndersonLealIsraelachvili2010_Langmuir}.

In
this work we focus on regimes where the ratio is of order unity, between
the weak adhesion (adhesion energy/bending energy $\ll 1$) and strong adhesion (adhesion energy/bending energy $\gg 1$)
regimes.  In this regime the membrane deformation may increase the
``contact area'' between two vesicles and enhance the adhesion effects.
The equilibration of vesicle membranes in the strong adhesion regime
have been well studied and documented
(see~\cite{RamachandranAndersonLealIsraelachvili2010_Langmuir,SteinkuhlerAgudo-Canalejo2016_BJ,FlormannAouane2017_SciReports} and
references therein). However it is unclear how the adhesion couples to
the vesicle hydrodynamics in such regime.  Our work in this paper aims
to address this question with quantitative characterization in terms of
physical parameters.

@@

In a quiescent environment, sub-micron size vesicles are found to form a
doublet due to their van der Waals' attractive
interactions~\cite{RamachandranAndersonLealIsraelachvili2010_Langmuir}.
Due to the strong van der Waals' adhesive force, the vesicles are far
from spherical shape and the membrane is almost flat in the ``contact"
region. Gires {\it et al.} used small-deformation analysis to
investigate the hydrodynamic interactions between two vesicles in a
planar shear flow with a long separation
distance~\cite{GiresDankerMisbah2012_PRE}. They found that the vesicle
interaction could be either attractive or repulsive depending on the
organization of the two vesicles relative to the shear flow. To the best
of authors' knowledge, the effects of close-range vesicle adhesion on
their hydrodynamics under an external flow have not been studied and
quantified. The goal of this work is to use state-of-the-art boundary
integral simulations to numerically investigate the dynamics of two
vesicles in both planar shear flow and extensional flow for a wide range
of vesicle shapes and adhesive strength and distance.

Boundary integral equation (BIE) approaches are well-suited for solving
the low Re flow problems considered here as they lead to reduction in
dimensionality and achieve high-order accuracy even for moving geometry
problems. When the vesicles adhere, one major issue for BIE solvers is
to resolve the vesicle-vesicle hydrodynamic forces which become {\em
nearly singular}. In this work, we use the recently developed
spectrally-accurate algorithm of~\cite{barnett2015spectrally} for
overcoming this problem. To overcome the numerical stiffness induced by
the membrane bending forces, we extend the adaptive time-stepping scheme
developed in~\cite{quaife2016adaptive} to our setting.

@@

\begin{figure}[htp]
  \includegraphics[width = 0.19\textwidth,trim={5cm 2cm 5cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi2em2_ra070_image01.png}
  \includegraphics[width = 0.19\textwidth,trim={5cm 2cm 5cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi2em2_ra070_image02.png}
  \includegraphics[width = 0.19\textwidth,trim={5cm 2cm 5cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi2em2_ra070_image03.png}
  \includegraphics[width = 0.19\textwidth,trim={5cm 2cm 5cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi2em2_ra070_image04.png}
  \includegraphics[width = 0.19\textwidth,trim={5cm 2cm 5cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi2em2_ra070_image05.png}
  \caption{\label{fig:extensional1} Two vesicles in an extensional flow.
  The reduced area is $A = 0.7$, the Hamaker constant is $H = 0.7$, and
  the adhesion length scale is $\delta = 0.4$, extensional rate is $\chi
  = 0.02$.  The center of the vesicle, denoted with a black square, is a
  stagnation point.}
\end{figure}

\begin{figure}[htp]
  \includegraphics[width = 0.19\textwidth,trim={5cm 2cm 5cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi7em2_ra070_image01.png}
  \includegraphics[width = 0.19\textwidth,trim={5cm 2cm 5cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi7em2_ra070_image02.png}
  \includegraphics[width = 0.19\textwidth,trim={5cm 2cm 5cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi7em2_ra070_image03.png}
  \includegraphics[width = 0.19\textwidth,trim={5cm 2cm 5cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi7em2_ra070_image04.png}
  \includegraphics[width = 0.19\textwidth,trim={5cm 2cm 5cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi7em2_ra070_image05.png}
  \caption{\label{fig:extensional2} Two vesicles in an extensional flow.
  The reduced area is $A = 0.7$, the Hamaker constant is $H = 0.7$, and
  the adhesion length scale is $\delta = 0.4$, extensional rate is $\chi
  = 0.07$.  The center of the vesicle, denoted with a black square, is a
  stagnation point.}
\end{figure}

\begin{figure}[htp]
  \includegraphics[width = 0.19\textwidth,trim={5cm 2cm 5cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi1em1_ra070_image01.png}
  \includegraphics[width = 0.19\textwidth,trim={5cm 2cm 5cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi1em1_ra070_image02.png}
  \includegraphics[width = 0.19\textwidth,trim={5cm 2cm 5cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi1em1_ra070_image03.png}
  \includegraphics[width = 0.19\textwidth,trim={5cm 2cm 5cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi1em1_ra070_image04.png}
  \includegraphics[width = 0.19\textwidth,trim={5cm 2cm 5cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi1em1_ra070_image05.png}
  \caption{\label{fig:extensional4} Two vesicles in an extensional flow.
  The reduced area is $A = 0.7$, the Hamaker constant is $H = 0.7$, and
  the adhesion length scale is $\delta = 0.4$, extensional rate is $\chi
  = 0.1$.  The center of the vesicle, denoted with a black square, is a
  stagnation point.}
\end{figure}

\begin{figure}[htp]
  \includegraphics[width = 0.19\textwidth,trim={5cm 2cm 5cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi1em1_ra080_image01.png}
  \includegraphics[width = 0.19\textwidth,trim={5cm 2cm 5cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi1em1_ra080_image02.png}
  \includegraphics[width = 0.19\textwidth,trim={5cm 2cm 5cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi1em1_ra080_image03.png}
  \includegraphics[width = 0.19\textwidth,trim={5cm 2cm 5cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi1em1_ra080_image04.png}
  \includegraphics[width = 0.19\textwidth,trim={5cm 2cm 5cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi1em1_ra080_image05.png}
  \caption{\label{fig:extensional3} Two vesicles in an extensional flow.
  The reduced area is $A = 0.8$, the Hamaker constant is $H = 0.7$, and
  the adhesion length scale is $\delta = 0.4$, extensional rate is $\chi
  = 0.1$.  Initially, the doublet tilts to an inclination angle, and
  then the extensional rate overcomes the adhesive force and the doublet
  is broken.  The center of the vesicle, denoted with a black square, is
  a stagnation point.}
\end{figure}

We consider six pairs of vesicles with reduced areas 0.70, 0.75, 0.80,
0.85, 0.90, and 0.95, and we vary the extensional rate, $\chi$, between
$10^{-2}$ and $10^{-1}$.  The length of each of the vesicles is fixed,
and the area is changed to achieve the desired reduced area.  Since the
stagnation point can be controlled in an experimental
setting~\cite{Johnson-Chavarria2011_EMJ}, we translate the vesicles at
each time step so that the stagnation point occurs exactly in the middle
of the two vesicles.  With this adjustment, the vesicles either remain
as a doublet in the fluid trap centered around the stagnation point, or
the doublet is broken and the vesicles diverge apart from one another.

>>>>

\title{Hydrodynamics and rheology of adhesive vesicle doublets}
% Force line breaks with \\
%\thanks{A footnote to the article title}%

@@

\section{Introduction}
Vesicles (closed fluid-filled phospholipid bilayer membranes) have been
widely utilized as biological cell mimics in biophysics and material
engineering \cite{sackmann1996}.  The ever expanding applications have
encouraged detailed experimental, theoretical, and numerical
investigations of vesicle hydrodynamics in applied flow, electric
fields, and magnetic fields.  Theoretical investigations often take the
form of small-deformation analysis or spheroidal dynamics of a
nearly-spherical vesicle (or nearly-circular in two dimensions)
subjected to an applied flow or under an external
forcing~\cite{Barthes-BieselRallison1981_JFM, Misbah2006_PRL,
Vlahovska2007_PRE, Finken2008_EPL, ZhangZahnTanLin2013_PoF,
Nganguia2013_PRE}.  On the other hand, various numerical methods have
been developed for simulating the transient hydrodynamics of vesicle
suspensions~\cite{BagchiJohoson2005_JBE, Biben2005_EJP,
Veerapaneni2009_JCP, SeolHuKimLai2016_JCP}.

Hydrodynamics of a single vesicle in Stokes flow has been extensively
investigated. In a planar shear flow, the vesicle hydrodynamics is
characterized by the reduced volume (reduced area in two dimensions),
viscosity contrast between interior and exterior fluids, and shear rate
of the imposed far-field fluid flow. In addition, a vesicle with a rigid
particle inside is also investigated as a biological mimic of a
eukaryotic cell with a nucleus that occupies nearly half of the
intracellular volume~\cite{Veerapaneni2011_PRL}. Small-deformation
analysis shows that a vesicle tank-treads in a planar shear flow for low
viscosity contrast and shear rate. At high viscosity contrast the
tank-tread (TT) dynamics transitions to tumbling (TB)
dynamics~\cite{Misbah2006_PRL, Vlahovska2007_PRE}, and this leads to a
transition in effective shear viscosity in the vesicle
suspension~\cite{Misbah2006_PRL,Vitkova2008_BJ} that is also validated
by direct numerical simulations~\cite{GhigliottiBibenMisbah2010_JFM} and
experiments~\cite{KantslerSegreSteinberg2008_EPL,
ZabuskySegreDeschamps2011_PoF}.  Between TT and TB vesicle dynamics, a
breathing (tremble) mode is also
observed~\cite{Misbah2006_PRL,KantslerSegreSteinberg2008_PRL,
ZhaoShaqfeh2011_JFM, SpannZhaoShaqfeh2014_PoF} to affect the effective
shear viscosity in a different way.  In an extensional flow (planar or
uniaxial), vesicle shape dynamics depends sensitively on the vesicle
reduced volume and the elastic capillary
number~\cite{KantslerSegreSteinberg2008_PRL, ZhaoShaqfeh2013_JFM,
Narsimhan2014_JFM, DahlNarsimhanGouveia2016_SoftMatt}: Asymmetric shape
and oscillatory undulation of the vesicle membrane are two examples of
the complex vesicle hydrodynamics in an extensional flow.

Collective hydrodynamics of vesicles is dictated by the vesicle-vesicle
interactions.  In a quiescent flow, vesicle-vesicle adhesion leads to
the formation of vesicle doublets
\cite{Ziherl2007_PRL,ZiherlSvetina2007_PNAS} or clusters
\cite{SvetinaZiherl2008_Bioelectrochemistry,FlormannAouane2017_SciReports}.
As a model for red blood cell (RBC) aggregates, an idealized model for
adhesive vesicle-vesicle interactions gives rise to vesicle
configurations similar to those observed in experiments of
fibrinogen-induced RBC
aggregates~\cite{SvetinaZiherl2008_Bioelectrochemistry,FlormannAouane2017_SciReports}.
Using the Lennard-Jones (L-J) potential for point-point interaction
between two vesicle membranes without the molecular details of adhesive
interactions between RBCs, Flormann {\em et al.}~found the contact zone
to buckle under strong adhesion~\cite{FlormannAouane2017_SciReports}.  A
direct consequence of the buckling is the sigmoidal shape of the
adhesion region, which is also observed in RBC doublets and explained in
a slightly different model~\cite{ZiherlSvetina2007_PNAS}.  It remains
unknown how adhesive interactions between vesicles might lead to
different vesicle hydrodynamics that results in different rheological
properties of a vesicle suspension. Such studies will provide useful
insight to the rheology of blood flow and how to use nano solutes to
control the hypertension by adjusting the adhesive interactions between
RBCs.

In numerical simulations of  a vesicle suspension, the adhesive
interaction between two vesicles is often ignored
\cite{Veerapaneni2009_JCP,RahimianVeerapaneniBiros2010_JCP} for
numerical convenience, mostly because a smaller time step is often
needed to resolve the dynamics of the lubrication thin film between two
membranes with adhesion than without.  However, adhesive interactions
between RBCs lead to different RBC clusters that are expected to alter
the rheological properties of the RBC
suspension~\cite{NeuMeiselman2002_BJ,
SvetinaZiherl2008_Bioelectrochemistry}. Hydrodynamic simulations of
adhesive vesicles in a quiescent flow have revealed physical insights to
the equilibrium shapes of RBC aggregates in both
experiments~\cite{FlormannAouane2017_SciReports} and
theory~\cite{ZiherlSvetina2007_PNAS}. The main goal of this work is to
investigate the hydrodynamics of adhesive vesicles in flowing conditions
common in microfluidic vesicle flow, such as a planar shear flow and an
extensional flow.

The electrostatic nature of lipid molecules (a hydrophobic tail and a
hydrophilic head with an electric dipole) complicates the interactions
between a lipid bilayer membrane and another bilayer
membrane~\cite{EvansMetcalfe1984_BJ, Book_PhysicalBasisCellAdhesion,
Book_IntermolecularSurfaceForces, PerutkovaFrank-Bertoncelij2013_CSB} or
a solid (such as a glass substrate or a nano particle).  Adhesion
between a vesicle and a substrate has been extensively studied with more
focuses on the static equilibrium shapes~\cite{Seifert1990_PRA,
ShiFengGao2006_ActaMechSin, LinFreund2007_IntJSolidsStructures,
GruhnFrankeDimova2007_Langmuir,das2008adhesion, zhang2009phase,
SteinkuhlerAgudo-Canalejo2016_BJ} than the transient adhesion
process~\cite{cantat1999lift, suk-sei2001, BlountMiksisDavis2013_PRSa}.
Adhesive interactions between lipid membranes are essential in many
biomedical, biological, and biophysical processes.  For example, vesicle
adhesion is crucial to initiate membrane fusion and fission in
endocytosis and exocytosis, the transport of small vesicles through
membrane surfaces.  In the absence of an external electric field and
ions in the solvents, it is reported that the adhesive interactions
between two membranes can be well approximated by the L-J-type
potential~\cite{FlormannAouane2017_SciReports}, that consists of a
long-range attraction component and a short-range repulsion
component~\cite{Book_IntermolecularSurfaceForces}, and the combination
of the two gives rise to adhesion at the separation distance that
minimizes the interaction
potential~\cite{Book_IntermolecularSurfaceForces}.

Evans and Metcalfe~\cite{EvansMetcalfe1984_BJ} used the micropipette to
measure the adhesive force between two vesicles bound by strong
adhesion.  They were able to measure the reduction in the free energy
per unit area of membrane-membrane contact formation due to the van der
Waals' attraction. For the adhesive interaction between two lipid
bilayer membranes in buffer solutions, a typical range for the energy
density is between $1$ to $10$ $\mu J/m^2$, similar to the adhesion
energy density of $3$ $\mu J/m^2$ between two
RBCs~\cite{FlormannAouane2017_SciReports}.  For adhesion interaction
between a vesicle and a substrate~\cite{GruhnFrankeDimova2007_Langmuir},
the boundary between weak and strong adhesion is around $1$ $\mu J/m^2$:
the vesicle-substrate interaction is ``strong" when the adhesion energy
density is larger than $1$ $\mu J/m^2$ and ``weak" when the energy
density is less than $1$ $\mu J/m^2$.  In this work we use the ratio of
the adhesion energy to the bending energy for classification of adhesion
strength. The ratio of total adhesion energy to the bending energy
(bending rigidity of lipid bilayer membranes $\sim 10^{-19}$ $J$) gives
a measure of vesicle deformation in the presence of
adhesion~\cite{RamachandranAndersonLealIsraelachvili2010_Langmuir}.

In this work we focus on regimes where the ratio is of order one,
between the weak adhesion (adhesion energy/bending energy $\ll 1$) and
strong adhesion (adhesion energy/bending energy $\gg 1$) regimes.  In
this regime the membrane deformation may increase the ``contact area''
between two vesicles and enhance the adhesion effects.  The
equilibration of vesicle membranes in the strong adhesion regime have
been well-studied and documented
(see~\cite{RamachandranAndersonLealIsraelachvili2010_Langmuir,
SteinkuhlerAgudo-Canalejo2016_BJ, FlormannAouane2017_SciReports} and
references therein). However it is unclear how the adhesion couples to
the vesicle hydrodynamics in this regime.  This paper aims to address
this question with quantitative characterization in terms of physical
parameters.

@@

In a quiescent environment, sub-micron size vesicles are found to form a
doublet due to their van der Waals' attractive
interactions~\cite{RamachandranAndersonLealIsraelachvili2010_Langmuir}.
Due to the strong van der Waals' adhesive force, the vesicles are far
from spherical shape and the membrane is almost flat in the ``contact"
region. Gires {\em et al.}~used small-deformation analysis to
investigate the hydrodynamic interactions between two vesicles in a
planar shear flow with a long separation
distance~\cite{GiresDankerMisbah2012_PRE}. They found that the vesicle
interaction could be either attractive or repulsive depending on the
organization of the two vesicles relative to the shear flow. To the best
of authors' knowledge, the effects of close-range vesicle adhesion on
their hydrodynamics under an external flow have not been studied and
quantified. The goal of this work is to use state-of-the-art boundary
integral simulations to numerically investigate the dynamics of two
vesicles in both planar shear flow and extensional flow for a wide range
of vesicle shapes and adhesive strength and distance.

Boundary integral equation (BIE) approaches are well-suited for solving
the low Reynodls flow problems considered here as they lead to reduction
in dimensionality and achieve high-order accuracy even for moving
geometry problems. The BIE resolves the self-vesicle interaction using
quadrature for logarithmically singular integrands~\cite{alp1999}, but
as the vesicles adhere, the BIE solver must also resolve the
vesicle-vesicle {\em nearly-singular} hydrodynamic forces.  We use an
interpolation-based quadrature rule~\cite{qua-bir2014} to maintain
high-order accuracy for all vesicle-vesicle separation distances.  To
overcome the numerical stiffness induced by the membrane bending forces
and to control the global error, we employ a high-order spectrally
deferred implicit-explicit adaptive time stepping
scheme~\cite{quaife2016adaptive}.

@@

\begin{figure}[htp]
  \includegraphics[width = 0.19\textwidth,trim={5cm 2cm 5cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi2em2_ra070_image01.png}
  \includegraphics[width = 0.19\textwidth,trim={5cm 2cm 5cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi2em2_ra070_image02.png}
  \includegraphics[width = 0.19\textwidth,trim={5cm 2cm 5cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi2em2_ra070_image03.png}
  \includegraphics[width = 0.19\textwidth,trim={5cm 2cm 5cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi2em2_ra070_image04.png}
  \includegraphics[width = 0.19\textwidth,trim={5cm 2cm 5cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi2em2_ra070_image05.png}
  \caption{\label{fig:extensional1} Two vesicles in an extensional flow.
  The reduced area is $A = 0.7$, the Hamaker constant is $H = 0.7$, the
  adhesion length scale is $\delta = 0.4$, and the extensional rate is
  $\chi = 0.02$.  The center of the vesicle, denoted with a black
  square, is a stagnation point.}
\end{figure}

\begin{figure}[htp]
  \includegraphics[width = 0.19\textwidth,trim={5cm 2cm 5cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi7em2_ra070_image01.png}
  \includegraphics[width = 0.19\textwidth,trim={5cm 2cm 5cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi7em2_ra070_image02.png}
  \includegraphics[width = 0.19\textwidth,trim={5cm 2cm 5cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi7em2_ra070_image03.png}
  \includegraphics[width = 0.19\textwidth,trim={5cm 2cm 5cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi7em2_ra070_image04.png}
  \includegraphics[width = 0.19\textwidth,trim={5cm 2cm 5cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi7em2_ra070_image05.png}
  \caption{\label{fig:extensional2} Two vesicles in an extensional flow.
  The reduced area is $A = 0.7$, the Hamaker constant is $H = 0.7$, the
  adhesion length scale is $\delta = 0.4$, and the extensional rate is
  $\chi = 0.07$.  The center of the vesicle, denoted with a black
  square, is a stagnation point.}
\end{figure}

\begin{figure}[htp]
  \includegraphics[width = 0.19\textwidth,trim={5cm 2cm 5cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi1em1_ra070_image01.png}
  \includegraphics[width = 0.19\textwidth,trim={5cm 2cm 5cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi1em1_ra070_image02.png}
  \includegraphics[width = 0.19\textwidth,trim={5cm 2cm 5cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi1em1_ra070_image03.png}
  \includegraphics[width = 0.19\textwidth,trim={5cm 2cm 5cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi1em1_ra070_image04.png}
  \includegraphics[width = 0.19\textwidth,trim={5cm 2cm 5cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi1em1_ra070_image05.png}
  \caption{\label{fig:extensional3} Two vesicles in an extensional flow.
  The reduced area is $A = 0.7$, the Hamaker constant is $H = 0.7$, the
  adhesion length scale is $\delta = 0.4$, and the extensional rate is
  $\chi = 0.1$.  The center of the vesicle, denoted with a black square,
  is a stagnation point.}
\end{figure}

\begin{figure}[htp]
  \includegraphics[width = 0.19\textwidth,trim={5cm 2cm 5cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi1em1_ra080_image01.png}
  \includegraphics[width = 0.19\textwidth,trim={5cm 2cm 5cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi1em1_ra080_image02.png}
  \includegraphics[width = 0.19\textwidth,trim={5cm 2cm 5cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi1em1_ra080_image03.png}
  \includegraphics[width = 0.19\textwidth,trim={5cm 2cm 5cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi1em1_ra080_image04.png}
  \includegraphics[width = 0.19\textwidth,trim={5cm 2cm 5cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi1em1_ra080_image05.png}
  \caption{\label{fig:extensional4} Two vesicles in an extensional flow.
  The reduced area is $A = 0.8$, the Hamaker constant is $H = 0.7$, the
  adhesion length scale is $\delta = 0.4$, and the extensional rate is
  $\chi = 0.1$.  Initially, the doublet tilts to an inclination angle,
  and then the extensional rate overcomes the adhesive force and the
  doublet is broken.  The center of the vesicle, denoted with a black
  square, is a stagnation point.}
\end{figure}

We consider six pairs of vesicles with reduced areas 0.70, 0.75, 0.80,
0.85, 0.90, and 0.95, and we vary the extensional rate, $\chi$, between
$10^{-2}$ and $10^{-1}$.  The length of each of the vesicles is fixed,
and the area is changed to achieve the desired reduced area.  Since the
stagnation point can be controlled in an experimental
setting~\cite{Johnson-Chavarria2011_EMJ}, we translate the center of the
doublet at each time step so that the stagnation point occurs exactly in
the middle of the doublet.  With this adjustment, the vesicles either
remain as a doublet in the fluid trap centered around the stagnation
point, or the doublet is broken and the vesicles diverge apart from one
another.
